---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "XXXX"
  email: "XXXX@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---
16/20.

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
table(Cars93$Type)
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
typevec = (Cars93$Type)
barplot(table(Cars93$Type),xlab = "Types", ylab="Frequency", main = "Types Frequency", cex.lab= 1.3, cex.axis = 1.2, col=c(2,3,4))

```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(table(Cars93$Type), main = "Car Types", cex.lab = 1.2, cex.axis = 1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.

#Midsize

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist(Cars93$Horsepower, cex.lab = 1.2, cex.axis = 1.2, col = 3, main = "Horsepower Frequency", xlab = "Horsepower")
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist(Cars93$Horsepower, cex.lab = 1.2, cex.axis = 1.2, col = 3, breaks = 12, main = "Horsepower Frequency", xlab = "Horsepower")
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
# The second histogram allows for better visualization in terms of more specific sample groups being represented rather than seeing them in larger groups

# Because the 2nd histogram has classes approximately equal to $\sqrt{n=93}$ (-2)
```{r}

```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
# This type of variable can be classified as a categorical nominal variable since it represents distinct categories
```{r}
hist(Cars93$Passengers, cex.lab = 1.2, cex.axis = 1.2, col = 3, main = "Passengers Frequency", xlab = "Passengers")

```
# The previous histogram is mainly used for continuous numerical variables and soes not provide a good visualization for the data at hand
b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)

```{r}
barplot(table(Cars93$Passengers),xlab = "Passengers", ylab="Frequency", main = "Passengers Frequency", cex.lab= 1.3, cex.axis = 1.2, col= 3)

# This bar plot should show the frequency of cars that carry 1 and 3 passengers as 0 (-2)
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist(Cars93$Length, cex.lab = 1.2, cex.axis = 1.2, col = 3, main = "Length Frequency", xlab = "Length")
```

b) What is the shape of the distribution?
#The shpae is unimodal with a slight right skew

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
mean(Cars93$Length)
median(Cars93$Length)
```
# These do meet my expectations. The mean is slightly affected by extreme values in comparison to the median
You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.


