---
title: "HW 7"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Justin Worobel"
  email: "worobelj@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
editor_options: 
  chunk_output_type: console
---
18/20. Code breaking. 
**Due date:** December 4 beginning of class.


```{r echo=FALSE}
# Custom options for knitting
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  fig.align = "center",
  cache = FALSE
) 
```


In this homework, we will analyze news headlines data scraped from abcnews, posted on Github in a csv file named `news.csv`.

Read the `news.csv` into R and create the object `news` using

```{r}
library(tidyverse)
library(tm)
news<-read.csv("news.csv",header=T)

```

Read also the positive and negative word dictionaries, both found on Github. This will later come in handy:

```{r}
posWords <- scan("positive-words.txt", character(0), sep = "\n")  # 2006 items
negWords <- scan("negative-words.txt", character(0), sep = "\n")  # 4783 items
head(posWords,15)
head(negWords,15)
```


# Question 1
First, check how many years of news data we have loaded in in R.
```{r}
years = length(unique(news$year))
cat(years, "years")
```

# Question 2
Store the headlines column in a vector of text named `charVector`. Print the first 6 entries in the vector.
```{r}
charVector = news$headline_text
print(head(charVector, 6))
```

# Question 3
Convert `charVector` into a vector source object named `wordVector` using the function `VectorSource()`. (See class notes on how to do this) Then convert `wordVector` into a corpus object named `wordCorpus` using the function `Corpus()`.
```{r}
wordVector = VectorSource(charVector)
wordCorpus = Corpus(wordVector)
```

# Question 4
We are now ready to do some basic trimming on the text. Do the following on the `wordCorpus` object using the `tm_map()` function, and print what's left of the first paragraph after trimming. (Note: ignore warnings.)
* Make all text lowercase
* Remove punctuation
* Remove numbers
* Remove stopwords (e.g., the, a, at, etc.)

```{r}
#function for transformations 
transform = function(doc) {
  doc = tolower(doc)
  doc = removePunctuation(doc)
  doc = removeNumbers(doc)
  doc = removeWords(doc, stopwords("en"))
  return(doc)
}
wordCorpus = tm_map(wordCorpus, content_transformer(transform))

print(content(wordCorpus[[1]]))
```

# Question 5
What is a term document matrix? Explain. Create a term document matrix from the news vector and store it as an object named `tdm` using the function `TermDocumentMatrix`.
```{r}
# A term document matrix is a representation of a text corpus. rows represent words and columns represents documents. Each can represent the frequency of a term in a document
textCorpus = Corpus(VectorSource(charVector))
textCorpus = tm_map(textCorpus, content_transformer(transform))
tdm = TermDocumentMatrix(textCorpus)
```

# Question 6
Convert `tdm` into a matrix named `m`. Store the frequency of occurrence of each word in the news headlines in an object named `wordCounts`. Return the top 10 most frequent words. (Note: if you run into memory issues creating the matrix `m`, try using one of the computers in Tome.)
```{r}
m = as.matrix(tdm)
wordCounts = rowSums(m)
wordCounts_df = data.frame(word = names(wordCounts), count = wordCounts, row.names = NULL)
top_words = head(wordCounts_df[order(-wordCounts_df$count),], 10)
print(top_words)
```

# Question 7
Create a barplot of the words that showed up at least 50 times in the news headlines. Rotate the x axis labels by 90 degrees, and decrease the label font size by 25%.

```{r}
library(ggplot2)
filtered_words = subset(wordCounts_df, count >= 50)
barplot_data = ggplot(filtered_words, aes(x = reorder(word, -count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top Words in Headlines (appear >=50 times)",
       x = "Words",
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(.75)))

print(barplot_data)

    

```


# Question 8
What is the percentage of positive words in the news headlines? What is the percentage of negative words? Make a barplot for each (for positive and negative words that showed up separately) showing the most frequent words for words that showed up at least 20 times. (Note: don't forget to get the total number of unique words that showed up in the data.)

```{r}
library(dplyr)

freq_words_20 <- wordCounts[wordCounts >= 20]

total_unique_words <- length(wordCounts)

positive_words <- intersect(names(freq_words_20), posWords)
percentage_positive <- length(positive_words) / total_unique_words * 100

negative_words <- intersect(names(freq_words_20), negWords)
percentage_negative <- length(negative_words) / total_unique_words * 100

cat("Percentage of Positive Words:", percentage_positive, "%\n")
cat("Percentage of Negative Words:", percentage_negative, "%\n")

barplot(freq_words_20[positive_words], las = 2, cex.names = 0.75, main = "Most Frequent Positive Words (>= 20 times)")

barplot(freq_words_20[negative_words], las = 2, cex.names = 0.75, main = "Most Frequent Negative Words (>= 20 times)")

```

Let's get the number of articles published in each year and month. Note that the chunk below creates two new variables `count` and `yearmonth`:

```{r}
news <- news %>% group_by(year,month) %>% mutate(count=n(), yearmonth = paste(year, month,sep = '/')) %>% arrange(year,month,day)
```

# Question 9
Using `ggplot2`, create a barplot for the frequency of articles released in each year and month. On the x axis, you should have year month, e.g., 2003/10, 2003/11, 2003/12, 2004/1, on the y-axis you should have the number of articles released in the data. Do you see a change in the number of articles released across years? (Hint: use `factor()` with `levels=` options specified when you call `aes(x=)` to have a proper chronological sorting on your x axis. Use `+theme(axis.text=element_text(size=4,angle=90)` to fit the dates on the x axis.)

```{r}

ggplot(news, aes(x = factor(yearmonth, levels = unique(yearmonth)), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Number of Articles Released by Year and Month",
       x = "Year Month",
       y = "Number of Articles Released") +
  theme(axis.text.x = element_text(size = 4, angle = 90)) +
  scale_x_discrete(labels = function(x) format(as.Date(x, "%Y/%m"), "%Y/%m"))
```

Let's now practice working with `quanteda` and `corpus` packages. Install the packages first if you don't have them in your computer and load them in in R.

```{r}
library("quanteda")
```

# Question 10
Using the `term_stats()` function, return the 20 most frequent single words that show up in the data. Make sure to filter out
* punctuation
* symbols
* stopwords

```{r}
library("quanteda")
my_corpus = corpus(charVector)

# Tokenize the corpus
my_tokens = tokens(my_corpus, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE, remove_twitter = TRUE, to_lower = TRUE)

# Create a document-feature matrix (dfm)
my_dfm = dfm(my_tokens)

# Get term frequencies
term_freqs = rowSums(my_dfm)

# Combine term frequencies with terms
term_stats = data.frame(term = names(term_freqs), frequency = term_freqs)

# Filter out stopwords, short words, etc.
filtered_term_stats = term_stats %>%
  filter(nchar(term) > 1)

# Get the top 20 words
top_words = head(filtered_term_stats[order(-filtered_term_stats$frequency),], 20)

print(top_words)

```

# Question 11
Repeat the previous question for word pairs (frequency of 2 words that show up, instead of 1 word)
```{r}
my_corpus = corpus(charVector)

# Tokenize the corpus
my_tokens = tokens(my_corpus, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE, remove_twitter = TRUE, to_lower = TRUE)

# Create bigrams
my_bigrams_list = tokens_ngrams(my_tokens, n = 2)
my_bigrams = unlist(my_bigrams_list)

# Calculate bigram frequencies
bigram_freqs = table(my_bigrams)

# Convert to a data frame
bigram_stats = data.frame(bigram = names(bigram_freqs), frequency = as.numeric(bigram_freqs))

# Filter out stopwords, short words, etc.
filtered_bigram_stats = bigram_stats %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(nchar(word1) > 1, nchar(word2) > 1)

# Get the top 20 bigrams
top_bigrams = head(filtered_bigram_stats[order(-filtered_bigram_stats$frequency), ], 20)

print(top_bigrams)
```

# Question 12
Create a corpus object using `corpus()` function. Store it in an object named `newscorpus`. Convert `newscorpus` into paragraphs with `corpus_reshape()`.
```{r}
newscorpus <- corpus(news$headline_text)

newscorpus <- corpus_reshape(newscorpus, to = "paragraphs")

head(newscorpus)
```

# Question 13
Create a term document matrix named `news_dtm` by using the `dfm()` function. While you call `dfm()`, make sure to clean the paragraphs by keeping only the stems of the words, removing punctuations, removing symbols, removing numbers, and removing stopwords. If you wish to get rid of anything that does not look correct in the data (such as weird bunch of characters that show up as "words") you may trim them out at this stage using `dfm_remove()`. Finally, trim `news_dtm` such that it contains only words that show up at least 50 times. Print the first 6 rows of `news_dtm`. (Note: Again, ignore warnings.)
```{r}
news_tokens <- tokens(news$headline_text, remove_punct = TRUE)

news_dtm <- dfm(news_tokens, stem=TRUE, remove_punct=TRUE, remove_symbols = TRUE, remove_numbers = TRUE, remove=c(stopwords("english")))

news_dtm <- dfm_trim(news_dtm, min_termfreq = 50)

print(news_dtm[1:6, ])
```

# Question 14
Create a wordcloud of `news_dtm` using `textplot_wordcloud()`. Comment on words that you were surprised and not surprised to see in the news headlines dataset.
```{r}
library("quanteda.textplots")
wordcloud <- textplot_wordcloud(news_dtm, min_count = 50, color = "blue", scale = c(3, 0.5))

print(wordcloud)
```


Let's now do topic modeling. Make sure to load necessary packages via

```{r}
library("topicmodels")
library('tidytext')
```

# Question 14
Run the LDA algorithm using k = 8. Store your output in object named `topic_model`. Print the 10 most frequent words that occurred in each 8 topic.

```{r}
library(dplyr)
news_topics <- convert(news_dtm, to = "topicmodels")
topic_model = LDA(news_topics, method = "VEM", k=8)
top_terms_beta <- tidy(topic_model, matrix = "beta") %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

ggplot(top_terms_beta, aes(x = reorder(term, beta), y = beta, fill = factor(topic))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 10 Words with Highest Beta for Each Topic",
       x = "Word",
       y = "Beta Value",
       fill = "Topic") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Question 15
Plot betas of top 10 words with highest beta for each 8 topic. Comment on 3-4 news headline groups in terms of their topic content.
```{r}
top_terms_beta <- tidy(topic_model, matrix = "beta") %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

ggplot(top_terms_beta, aes(x = reorder(term, beta), y = beta, fill = factor(topic))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 10 Words with Highest Beta for Each Topic",
       x = "Word",
       y = "Beta Value",
       fill = "Topic") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Question 16
Plot gammas of top 5 documents with highest gamma for each 8 topic. Return the contents of the 5 documents with highest gamma only for topics 1 and 2. Does it makes sense that these documents are assigned to topics 1 and 2 accordingly? Comment.
```{r}
top_docs_gamma <- tidy(topic_model, matrix = "gamma") %>%
  group_by(topic) %>%
  top_n(5, gamma) %>%
  ungroup() %>%
  arrange(topic, -gamma)

ggplot(top_docs_gamma, aes(x = reorder(document, gamma), y = gamma, fill = factor(topic))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Top 5 Documents with Highest Gamma for Each Topic",
       x = "Document",
       y = "Gamma Value",
       fill = "Topic") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

top_docs_content <- news_dtm[top_docs_gamma$document, ]

print(top_docs_content)
print(top_docs_content)
```


You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


